---
title: "Social Disorder and Political Mobilisation"
subtitle: "SIS 750 - Data Analysis"
author: "Bern DySart"
institute: "American University"
date: "March 27, 2024"
output:
  xaringan::moon_reader:
    lib_dir: docs
    css: [default, shinobi, rladies-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r SetupChunk, echo = FALSE, warning = FALSE, include = FALSE}
library(tidyverse)
library(haven)
library(patchwork)
library(xaringanthemer)
library(forcats)

FctWhen = function(...){
  args = rlang::list2(...)
  rhs = map(args, rlang::f_rhs)
  cases = case_when(!!!args)
  exec(fct_relevel, cases, !!!rhs)
}

style_xaringan(
  text_font_size = "12rem"
  )

USD <- read_dta("events.dta")

USD2 = USD

USD = 
  USD |>
  mutate(
    TYPE = FctWhen(
      PTYPE == 10 ~ "General Warfare",
      PTYPE == 20 ~ "Inter-communal Warfare",
      PTYPE == 30 ~ "Armed Battle/Clash",
      PTYPE == 31 ~ "Armed Attack",
      PTYPE == 40 ~ "Pro-Government Terrorism (Repression)",
      PTYPE == 41 ~ "Anti-Government Terrorism",
      PTYPE == 42 ~ "Communal Terrorism",
      PTYPE == 50 ~ "Organised Violent Riot",
      PTYPE == 51 ~ "Spontaneous Violent Riot",
      PTYPE == 60 ~ "Organised Demonstration",
      PTYPE == 61 ~ "Pro-Government Demonstration",
      PTYPE == 62 ~ "Spontaneous Demonstration",
      PTYPE == 70 ~ "Other"
    )
  )

USD2 =
  USD |>
  summarise(
    mean = mean(TARGET1 == "Government"),
  ) |>
  mutate(
    REGION = "Global"
  )
```
# Data Introductory Information
## Urban Social Disorder Data version 3
- Event focused data set 
  - International data collection
  - Covering the years 1960-2014
      - 186 international capitals. 
  - Social Disorder and political mobilisation. 
- 13,460 total observations
  - 33 total variables
      - First and Secondary Target
      - First, Secondary, and Tertiary Actor
      - Type of Event
      - Length of Event
      - Number of Deaths
      - Number of Participants
- Citation:
  -   Thomson, Henry, Karim Bahgat, Henrik Urdal & Halvard Buhaug 2022. "Urban Social Disorder 3.0: A global, city-level event dataset of political mobilization and disorder." Journal of Peace Research, forthcoming
---
# Frequency of Event Types

- Internationally, what is the most frequent type of social disorder or mobilisation event?
  - The following graph spans the entire data set and all 13,460 observations
```{r ConflictChart, echo = FALSE, include = TRUE, fig.retina = 4, fig.align='left'}
USD |>  
  ggplot(aes(x = fct_infreq(TYPE), fill = TYPE))+
  coord_flip(ylim = c(0,5000)) + 
  geom_bar(show.legend = FALSE) +
  scale_y_continuous(
    expand = c(0,0),
    labels = scales::percent) +
  geom_text(
    stat = 'count', 
    aes(label = after_stat(round(count/sum(count)*100, digits = 1)), hjust = -0.5)
  ) +
  theme_minimal(base_size = 15) +
  theme(
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line = element_line(colour = "black"),
    panel.background = element_blank(), 
    panel.grid = element_blank()
  ) +
  labs(
    y = "Percentage"
  )
```
---
# Primary Target: Government

- What was the average amount of events that primarily targeted the government?
  - The graph shows the average of events targeting the government, both internationally and by individual region, between 1960 and 2014
```{r TargetAvg, echo = FALSE, include = TRUE, fig.align = 'center', fig.retina = 4}
USD |>
  group_by(REGION) |>
  summarise(
    mean = mean(TARGET1 == "Government")
  ) |>
  bind_rows(USD2) |> #adds the created variable that acts as the global sample mean
  ggplot(aes(
    x = fct_rev(fct_infreq(REGION, mean)), 
    y = mean,
    fill = (REGION == "Global")
  )) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = round(mean*100, digits = 1)), hjust = -0.5) +
  coord_flip() +
  scale_y_continuous(
    expand = c(0,0),
    limits = c(0,0.5)
  ) +
  theme_minimal(base_size = 15) +
  theme(
    axis.line = element_line(colour = "black"),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    panel.background = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(
    y = "Average"
  )